---
title: "2. Mucin_Expression_Processing"
author: "Baptiste Oosterlinck"
date: "21-2-2022"
output: html_document
---

#preparing the data frame
##1. Loading the metadata data into a dataframe
  - setting the dates into the correct format.
  - If you see numbers in stead of the correct dates go into excel:
    Data -> Text to Columns -> unselect all delimiters -> set column format to text
    
```{r}
ClinGC <- read_excel( path = meta_file, sheet = 1)
MucinGC <- read_excel( path = meta_file, sheet = 2)

#transform DOD, LFU, SD to date format in R
ClinGC$Sampling.Date <- as.Date(ClinGC$Sampling.Date , format = "%d/%m/%Y")
ClinGC$Date.of.Death <- as.Date(ClinGC$Date.of.Death , format = "%d/%m/%Y")
ClinGC$Last.Follow.Up <- as.Date(ClinGC$Last.Follow.Up , format = "%d/%m/%Y")

#mutate selected columns to factors
factor_cols <- c("Sample.Origin", "Tissue.Type", "Tumor.Location", "Lauren.Classification", "Survival", "Gender")
ClinGC[factor_cols] <- lapply(ClinGC[factor_cols], factor)

rm(factor_cols)
```

##2. Calculating the number of observational days using the sampling date - date of death or last follow up
```{r}
#calculating the number of observation days
ClinGC$Observation.Days <- rep(NA, nrow(ClinGC))

for (i in 1:nrow(ClinGC)) {
    if (is.na(ClinGC$Sampling.Date[i])) {

    } else if (!is.na(ClinGC$Date.of.Death[i])) {
        ClinGC$Observation.Days[i] <- ClinGC$Date.of.Death[i] - ClinGC$Sampling.Date[i]
    } else if (is.na(ClinGC$Date.of.Death[i]) & !is.na(ClinGC$Last.Follow.Up[i])) {
        ClinGC$Observation.Days[i] <- ClinGC$Last.Follow.Up[i] - ClinGC$Sampling.Date[i]
    }
}

rm(i)
```

##3. Mergin the datasets with the mucin expression data and the clinical data
```{r}
#merge both datasets
MetaData <- full_join(ClinGC,MucinGC, by = "Sample.ID") %>%
  as.data.frame(.)

rm(ClinGC,MucinGC)
```

##4. Transforming the stages of the TNM-staging from code to the real stage and determining the final cancer stage
```{r}
#binning of T1,T2 and T4 subclasses
T_stage <- c()
for (Stage in MetaData$T) {
    if (is.na(Stage)) {
        T_stage <- c(T_stage, NA)
    } else if (Stage == 1) {
        T_stage <- c(T_stage, "TX")
    } else if (Stage == 2) {
        T_stage <- c(T_stage, "T0")
    } else if (Stage == 3) {
        T_stage <- c(T_stage, "Tis")
    } else if (Stage == 4 | Stage == 5 | Stage == 6) {
        T_stage <- c(T_stage, "T1")
    } else if (Stage == 7 | Stage == 8 | Stage == 9) {
        T_stage <- c(T_stage, "T2")
    } else if (Stage == 10) {
        T_stage <- c(T_stage, "T3")
    } else if (Stage == 11 | Stage == 12) {
        T_stage <- c(T_stage, "T4a")
    } else if (Stage == 13) {
        T_stage <- c(T_stage, "T4b")
    }
}

N_stage <- c()
for (Stage in MetaData$N) {
    if (is.na(Stage)) {
        N_stage <- c(N_stage, NA)
    } else if (Stage == 1) {
        N_stage <- c(N_stage, "NX")
    } else if (Stage == 2) {
        N_stage <- c(N_stage, "N0")
    } else if (Stage == 3) {
        N_stage <- c(N_stage, "N1")
    } else if (Stage == 4 | Stage == 5 | Stage == 6) {
        N_stage <- c(N_stage, "N2")
    } else if (Stage == 7 | Stage == 8) {
        N_stage <- c(N_stage, "N3a")
    } else if (Stage == 9) {
        N_stage <- c(N_stage, "N3b")
    }
}

M_stage <- c()
for (Stage in MetaData$M) {
    if (is.na(Stage)) {
        M_stage <- c(M_stage, NA)
    } else if (Stage == 1) {
        M_stage <- c(M_stage, "MX")
    } else if (Stage == 2) {
        M_stage <- c(M_stage, "M1")
    } else if (Stage == 3) {
        M_stage <- c(M_stage, "M2")
    }
}

#adding the different stages to the metadata dataframe
MetaData$T <- T_stage
MetaData$N <- N_stage
MetaData$M <- M_stage
rm(T_stage, N_stage, M_stage,Stage)

#Determining tumor stage based upon TNM classification
CancerStage <- data.frame(Stage = rep(Inf, nrow(MetaData)))

StageMatrix <- t(matrix(c("1A", "1B", "2A", "2B", "3B", "2A",
                          "1B", "2A", "2B", "3A", "3B", "2B",
                          "2A", "2B", "3A", "3B", "3C", "3A",
                          "2B", "3A", "3A", "3B", "3C", "3A",
                          "3A", "3B", "3B", "3C", "3C", "3B",
                          "2A", "2B", "3A", "3B", "3C", "3A"
                          ), nrow = 6))

colnames(StageMatrix) <- c("N0", "N1", "N2", "N3a", "N3b", "NX")
rownames(StageMatrix) <- c("T1", "T2", "T3", "T4a", "T4b", "TX")
StageMatrix <- as.data.frame(StageMatrix)

for (i in 1:nrow(MetaData)) {
    if (!is.na(MetaData$M[i]) & MetaData$M[i] == "M1") {
        CancerStage[i, 1] <- 4
    } else if (is.na(MetaData$T[i]) | is.na(MetaData$N[i]) | is.na(MetaData$M[i])) {
        CancerStage[i, 1] <- NA
    } else {
        CancerStage[i, 1] <- StageMatrix[MetaData$T[i], MetaData$N[i]]
    }
}

MetaData$Stage <- as.factor(CancerStage$Stage)

rm(CancerStage, StageMatrix, i)
```

#General exploration of the mucin relative expressions

##1. Calculating the minimum, maximum and mean of each mucin for each tissue type
```{r}
Mucins <- c("MUC1", "MUC2", "MUC4", "MUC5AC", "MUC6", "MUC13")
Tissue <- levels(MetaData$Tissue.Type)

Mucin_metrics <- c()

for (i in Tissue) {
    temp_tissue <- MetaData[MetaData$Tissue.Type == i,]
    for (j in Mucins) {
        min_temp <- min(temp_tissue[, j], na.rm = TRUE)
        max_temp <- max(temp_tissue[, j], na.rm = TRUE)
        mean_temp <- mean(temp_tissue[, j], na.rm = TRUE)
        muc_type_temp <- paste(j, i, sep = "_")
        Mucin_metrics <- rbind(Mucin_metrics, c(muc_type_temp, min_temp, mean_temp, max_temp))
    }
    rm(min_temp, max_temp, mean_temp, muc_type_temp, temp_tissue)
}

rownames(Mucin_metrics) <- Mucin_metrics[, 1]
Mucin_metrics <- Mucin_metrics[, 2:4]
colnames(Mucin_metrics) <- c("min", "mean", "max")

rm(Tissue,i,j)
```

##2. Looking at the expression values of the control patient cohort:
    * Check normality of the distribution
    * caluclating confidence interval for each mucin
    
```{r}
#extracting the expression data from the control patients the metadata object
MetaData_control <- MetaData[MetaData$Tissue.Type == "noninflammed",]

#looking at the distribution of the functional dyspepsia patients
Control_Dens <- ggdensity(data = MetaData_control, x = "MUC13")
Control_Dens

#doing a log2 transformation and re-evaluating the normality
MetaData_control %>%
  mutate(.,MUC13_Log2 = log(MetaData_control$MUC13,2)) %>%
  ggdensity(., x = "MUC13_Log2")

#calculating the confidence intervals using a boot strapping method due to non-normal distribution
MUC_CI <- c()

for (MUC in Mucins){
  boot_MUC <- boot(data= MetaData_control[,MUC],function(x,i) mean(x[i]),R=10000)
  boot_MUC_CI <- boot.ci(boot_MUC,conf = 0.90,type = "bca")
  MUC_CI <- rbind(MUC_CI, boot_MUC_CI$bca[4:5])
  rm(boot_MUC,boot_MUC_CI,MUC)
}

row.names(MUC_CI) <- Mucins
colnames(MUC_CI) <- c("Lower_CI","Upper_CI")

#clean-up
rm(MetaData_control,Control_Dens)
```

##3. Per sample mucin expression level stratification based on CI of control samples

```{r}
#subsetting to data frame to the Adenocarinoma samples with their adjacent tissue samples
MUC_exp <- subset(MetaData, select = c("Patient.ID", "Tissue.Type", "MUC1", "MUC2", "MUC4", "MUC5AC", "MUC6", "MUC13"))
MUC_exp <- MUC_exp[MUC_exp$Tissue.Type != "noninflammed",] %>%
  drop_na(., "Tissue.Type")
#initializing the empty score matrix to hold the stratification scores
score_matrix_N <- as.data.frame(matrix(ncol = 7, nrow = 108))
colnames(score_matrix_N) <- c("Patient.ID", "MUC1_N", "MUC2_N", "MUC4_N", "MUC5AC_N", "MUC6_N", "MUC13_N")
score_matrix_T <- as.data.frame(matrix(ncol = 7, nrow = 108))
colnames(score_matrix_T) <- c("Patient.ID", "MUC1_T", "MUC2_T", "MUC4_T", "MUC5AC_T", "MUC6_T", "MUC13_T")
#getting the patient ID's ready in the score matrixes
score_matrix_N$Patient.ID <- MUC_exp$Patient.ID[MUC_exp$Tissue.Type == "adjacent"]
score_matrix_T$Patient.ID <- MUC_exp$Patient.ID[MUC_exp$Tissue.Type == "tumor"]

#j is for looping of the mucins
for (j in 3:8) {
    i_T <- 1
    i_N <- 1
#i is for looping of the patients (matrix rows)
    for (i in 1:nrow(MUC_exp)) {
        if (MUC_exp$Tissue.Type[i] == "adjacent") {

            s_N <- NA

            if (is.na(MUC_exp[i, j])) {

            } else if (MUC_exp[i, j] > MUC_CI[j - 2,2]) {
                s_N <- 1
            } else if (MUC_exp[i, j] > MUC_CI[j - 2,1]) {
                s_N <- 0
            } else {
                s_N <- -1
            }

            score_matrix_N[i_N, j - 1] <- s_N
            i_N <- i_N + 1

        } else if (MUC_exp$Tissue.Type[i] == "tumor") {

            s_T <- NA

            if (is.na(MUC_exp[i, j])) {

            } else if (MUC_exp[i, j] > MUC_CI[j - 2,2]) {
                s_T <- 1
            } else if (MUC_exp[i, j] > MUC_CI[j - 2,1]) {
                s_T <- 0
            } else {
                s_T <- -1
            }

            score_matrix_T[i_T, j - 1] <- s_T
            i_T <- i_T + 1

        }
    }
}

score_matrix <- full_join(score_matrix_N, score_matrix_T, by = "Patient.ID")
score_matrix <- score_matrix[, c(1, order(colnames(score_matrix)[2:13]) + 1)]

colnames_score_matrix <- colnames(score_matrix)
colnames_score_matrix <- c(colnames_score_matrix[1], paste("Strat.Score", colnames_score_matrix[2:13], sep = "_"))
colnames(score_matrix) <- colnames_score_matrix

MetaData <- full_join(MetaData, score_matrix, by = "Patient.ID")

rm(MUC_exp, s_N, s_T, score_matrix, score_matrix_N, score_matrix_T, i_N, i_T, i, j,colnames_score_matrix)
```

##4. Determining the Mucin phenotype Classification of each sample

```{r}
df_Strat.Score <- MetaData[, c("Patient.ID", "Tissue.Type", "Strat.Score_MUC1_N", "Strat.Score_MUC1_T", "Strat.Score_MUC2_N", "Strat.Score_MUC2_T", "Strat.Score_MUC4_N", "Strat.Score_MUC4_T", "Strat.Score_MUC5AC_N", "Strat.Score_MUC5AC_T", "Strat.Score_MUC6_N", "Strat.Score_MUC6_T", "Strat.Score_MUC13_N", "Strat.Score_MUC13_T")]

df_Strat.Score <- df_Strat.Score %>%
  mutate(., Strat.Score_MUC5AC_N = replace(Strat.Score_MUC5AC_N, Strat.Score_MUC5AC_N == 0, 1),
         Strat.Score_MUC5AC_T = replace(Strat.Score_MUC5AC_T, Strat.Score_MUC5AC_T == 0, 1),
         Strat.Score_MUC6_N = replace(Strat.Score_MUC6_N, Strat.Score_MUC6_N == 0, 1),
         Strat.Score_MUC6_T = replace(Strat.Score_MUC6_T, Strat.Score_MUC6_T == 0, 1),
         Strat.Score_MUC1_N = replace(Strat.Score_MUC1_N, Strat.Score_MUC1_N == 0, 1),
         Strat.Score_MUC1_T = replace(Strat.Score_MUC1_T, Strat.Score_MUC1_T == 0, 1)) %>%
  mutate(., Strat.Score_MUC5AC_N = replace(Strat.Score_MUC5AC_N, Strat.Score_MUC5AC_N == -1, 0),
         Strat.Score_MUC5AC_T = replace(Strat.Score_MUC5AC_T, Strat.Score_MUC5AC_T == -1, 0),
         Strat.Score_MUC6_N = replace(Strat.Score_MUC6_N, Strat.Score_MUC6_N == -1, 0),
         Strat.Score_MUC6_T = replace(Strat.Score_MUC6_T, Strat.Score_MUC6_T == -1, 0),
         Strat.Score_MUC1_N = replace(Strat.Score_MUC1_N, Strat.Score_MUC1_N == -1, 0),
         Strat.Score_MUC1_T = replace(Strat.Score_MUC1_T, Strat.Score_MUC1_T == -1, 0)) %>%
  mutate(., Strat.Score_MUC2_N = replace(Strat.Score_MUC2_N, Strat.Score_MUC2_N == -1, 0),
         Strat.Score_MUC2_T = replace(Strat.Score_MUC2_T, Strat.Score_MUC2_T == -1, 0),
         Strat.Score_MUC4_N = replace(Strat.Score_MUC4_N, Strat.Score_MUC4_N == -1, 0),
         Strat.Score_MUC4_T = replace(Strat.Score_MUC4_T, Strat.Score_MUC4_T == -1, 0),
         Strat.Score_MUC13_N = replace(Strat.Score_MUC13_N, Strat.Score_MUC13_N == -1, 0),
         Strat.Score_MUC13_T = replace(Strat.Score_MUC13_T, Strat.Score_MUC13_T == -1, 0))

criteria <- data.frame(Patient.ID = df_Strat.Score$Patient.ID, Tissue.Type = df_Strat.Score$Tissue.Type,
                       gast_MUC_N = df_Strat.Score$Strat.Score_MUC5AC_N + df_Strat.Score$Strat.Score_MUC6_N + df_Strat.Score$Strat.Score_MUC1_N,
                       gast_MUC_T = df_Strat.Score$Strat.Score_MUC5AC_T + df_Strat.Score$Strat.Score_MUC6_T + df_Strat.Score$Strat.Score_MUC1_T,
                       int_MUC_N = df_Strat.Score$Strat.Score_MUC13_N + df_Strat.Score$Strat.Score_MUC2_N + df_Strat.Score$Strat.Score_MUC4_N,
                       int_MUC_T = df_Strat.Score$Strat.Score_MUC13_T + df_Strat.Score$Strat.Score_MUC2_T + df_Strat.Score$Strat.Score_MUC4_T)
criteria <- criteria[criteria$Tissue.Type == "tumor", c(1, 3:6)]

mucine_type <- data.frame(Patient.ID = rep(NA, nrow(criteria)), Mucin.Phenotype = rep(NA, nrow(criteria)))

for (i in 1:nrow(criteria)) {
    mucine_type$Patient.ID[i] <- criteria$Patient.ID[i]
    if (is.na(criteria$gast_MUC_T[i]) | is.na(criteria$int_MUC_T[i])) {
        mucine_type$Mucin.Phenotype[i] <- NA
    } else if (criteria$gast_MUC_T[i] > 0 && criteria$int_MUC_T[i] > 0) {
        mucine_type$Mucin.Phenotype[i] <- "Mixed"
    } else if (criteria$gast_MUC_T[i] > 0 & criteria$int_MUC_T[i] == 0) {
        mucine_type$Mucin.Phenotype[i] <- "Gastric"
    } else if (criteria$gast_MUC_T[i] == 0 & criteria$int_MUC_T[i] > 0) {
        mucine_type$Mucin.Phenotype[i] <- "Intestinal"
    } else if (criteria$gast_MUC_T[i] == 0 & criteria$int_MUC_T[i] == 0) {
        mucine_type$Mucin.Phenotype[i] <- "Null"
    } else {
        mucine_type$Mucin.Phenotype[i] <- "not considered"
    }
}

mucine_type$Mucin.Phenotype <- as.factor(mucine_type$Mucin.Phenotype)
summary(mucine_type)

MetaData <- full_join(MetaData, mucine_type,by = "Patient.ID")

#cleanUp
rm("criteria","df_Strat.Score","Mucin_metrics","mucine_type", "i")
```

##5. Generating the boxplots for figure 1-A

```{r}
#creating long format for the mucin data
MetaData_long <- gather(data = MetaData , key = "Mucin", value = "CNRQ",  22:27) %>%
  mutate(CNRQ_log2 = log(x = .$CNRQ, base = 2))
#ordering the Mucin variable to have the gastric and intestinal mucins grouped
MetaData_long$Mucin <- factor(MetaData_long$Mucin, levels = c("MUC1","MUC5AC","MUC6","MUC2","MUC4","MUC13"), labels = c("MUC1","MUC5AC","MUC6","MUC2","MUC4","MUC13"),ordered = TRUE)
#replacing the noninflammed with control as tissue type
MetaData_long$Tissue.Type <- factor(MetaData_long$Tissue.Type,levels = c("noninflammed","adjacent","tumor"),labels = c("control","adjacent","tumor"),ordered = TRUE)

#Adding the expression levels of the tumours per mucin phenotype to the dataframe
MetaData_Extended <- MetaData_long %>%
  mutate(Mucin.Phenotype = .$Tissue.Type) %>%
  rbind(.,MetaData_long) %>%
  drop_na(., Mucin.Phenotype)

#generating general plot for the mucin expression: [control, adjacent, tumor] and [gastric, intestinal, mixed, null]
MucinGeneral_boxplot <- MetaData_Extended %>%
  ggplot(data = ., aes(x = Mucin.Phenotype , y = CNRQ_log2, color = Mucin.Phenotype))+
  geom_boxplot()+
  geom_point(position = position_jitter(seed = 1995,width = 0.15))+
  facet_wrap(~Mucin)+
  scale_x_discrete(guide = guide_axis(angle = 90))+
  stat_compare_means(aes(label = ..p.signif..),hide.ns = TRUE,method = "wilcox.test",comparisons = list(c("Gastric","Intestinal"),c("Intestinal","Mixed"),c("Mixed","Null"),c("Gastric","Mixed"),c("Gastric","Null"),c("Intestinal","Null")))+
  theme_classic2()
MucinGeneral_boxplot

rm(MetaData_Extended,MetaData_long)
```

#Correlation analysis using predictive power scores

```{r}
MetaData_PPS <- select(MetaData, c("Age","Gender","Tissue.Type","Tumor.Location","Lauren.Classification","T","N","M","G","Stage","Survival","Mucin.Phenotype","MUC1","MUC5AC","MUC6","MUC2","MUC4","MUC13"))

MetaData_PPS$Tumor.Location[MetaData_PPS$Tissue.Type == "adjacent"] <- NA
MetaData_PPS$Lauren.Classification[MetaData_PPS$Tissue.Type == "adjacent"] <- NA
MetaData_PPS$T[MetaData_PPS$Tissue.Type == "adjacent"] <- NA
MetaData_PPS$N[MetaData_PPS$Tissue.Type == "adjacent"] <- NA
MetaData_PPS$M[MetaData_PPS$Tissue.Type == "adjacent"] <- NA
MetaData_PPS$G[MetaData_PPS$Tissue.Type == "adjacent"] <- NA
MetaData_PPS$Stage[MetaData_PPS$Tissue.Type == "adjacent"] <- NA
MetaData_PPS$Mucin.Phenotype[MetaData_PPS$Tissue.Type == "adjacent"] <- NA

MetaData_PPS <- subset(MetaData_PPS, Tissue.Type != "noninflammed")
MetaData_PPS <- droplevels(MetaData_PPS,Tissue.Type = "noninflammed")

MetaData_PPS$T <- as.factor(MetaData_PPS$T)
MetaData_PPS$N <- as.factor(MetaData_PPS$N)
MetaData_PPS$M <- as.factor(MetaData_PPS$M)
MetaData_PPS$G <- as.factor(MetaData_PPS$G)
MetaData_PPS$Stage <- as.factor(MetaData_PPS$Stage)


PPS_df <- score_df(MetaData_PPS)
HeatMap_pps <- visualize_pps(MetaData_PPS,color_value_low = "white", color_value_high = "#7A0403", color_text = "black")
```

#Survival analysis
##1. Formatting the dataframe for survival analysis

```{r}
MetaData_Surv <- MetaData[MetaData$Tissue.Type == "tumor",]
MetaData_Surv$Survival <- as.numeric(MetaData_Surv$Survival)
MetaData_Surv <- drop_na(MetaData_Surv, Survival)


MetaData_Surv$Survival <- factor(MetaData_Surv$Survival)
MetaData_Surv$Survival <- as.numeric(MetaData_Surv$Survival)
MetaData_Surv <- drop_na(data = MetaData_Surv, "Survival")
MetaData_Surv$Mucin.Phenotype <- as.character(MetaData_Surv$Mucin.Phenotype)
```

##2. Kaplan- Meijer curve: mucin phenotype ~ survival (figure 2-A)

```{r}
##fitting a survival curve
SurvivalPhenotype <- survfit(Surv(Observation.Days, Survival) ~ Mucin.Phenotype, data = MetaData_Surv)
summary(SurvivalPhenotype)

phenoType_curve <- ggsurvplot(fit = SurvivalPhenotype, data = MetaData_Surv,
           pval = TRUE, 
           conf.int = TRUE, 
           risk.table = TRUE, 
           risk.table.col = "strata",
           surv.median.line = "hv",
           legend.labs = c("Gastric", "Intestinal", "Mixed", "Null"),
           break.x.by= 365,
           ggtheme = theme_classic())
phenoType_curve

rm(SurvivalPhenotype)
```

##3. Cox-proportional hazards model mucin phenotype (figure 2-A)
```{r}
#fitting a cox proportional hazards model
CoxPhenotype <- coxph(Surv(Observation.Days, Survival) ~ Age + Gender + Mucin.Phenotype, data = MetaData_Surv)
summary(CoxPhenotype)
Forest_coxPheno <- ggforest(CoxPhenotype)
Forest_coxPheno

#testing cox-proportional hazards model assumptions
test.pheno <- cox.zph(CoxPhenotype)
test.pheno

ggcoxzph(test.pheno)
rm(test.pheno)

ggcoxdiagnostics(CoxPhenotype,  type = "dfbeta", linear.predictions = FALSE, ggtheme = theme_bw())
ggcoxdiagnostics(CoxPhenotype,type = "deviance", linear.predictions = FALSE, ggtheme = theme_bw())
  
rm(CoxPhenotype)
```

##3 Kaplan-meijer curves and Cox-proportional hazards model per mucin (figure 2-B)
  For MUC1, MUC5AC and MUC6 the mid-level were filtered due to to low number of patient for the Kaplan-Meier curves.
  
```{r}
MetaData_Surv$Strat.Score_MUC13_T <- factor(MetaData_Surv$Strat.Score_MUC13_T, levels = c(-1,0,1), labels = c("low","mid","high"))
MetaData_Surv$Strat.Score_MUC1_T <- factor(MetaData_Surv$Strat.Score_MUC1_T, levels = c(-1,0,1), labels = c("low","mid","high"))
MetaData_Surv$Strat.Score_MUC2_T <- factor(MetaData_Surv$Strat.Score_MUC2_T, levels = c(-1,0,1), labels = c("low","mid","high"))
MetaData_Surv$Strat.Score_MUC4_T <- factor(MetaData_Surv$Strat.Score_MUC4_T, levels = c(-1,0,1), labels = c("low","mid","high"))
MetaData_Surv$Strat.Score_MUC5AC_T <- factor(MetaData_Surv$Strat.Score_MUC5AC_T, levels = c(-1,0,1), labels = c("low","mid","high"))
MetaData_Surv$Strat.Score_MUC6_T <- factor(MetaData_Surv$Strat.Score_MUC6_T, levels = c(-1,0,1), labels = c("low","mid","high"))

#Kaplan-Meijer curve
SurvivalMUC1 <- MetaData_Surv %>%
  filter(., Strat.Score_MUC1_T != "mid") %>%
  survfit(Surv(Observation.Days, Survival) ~ Strat.Score_MUC1_T, data = .)

MUC1_curve <- ggsurvplot(fit = SurvivalMUC1, data = MetaData_Surv, title = "MUC1", 
           pval = TRUE, 
           conf.int = TRUE, 
           risk.table = TRUE, 
           risk.table.col = "strata",
           surv.median.line = "hv",
           legend.labs = c("low","high"),
           break.x.by= 365,
           ggtheme = theme_classic()
           )
MUC1_curve
rm(SurvivalMUC1)

SurvivalMUC5AC <- MetaData_Surv %>%
  filter(., Strat.Score_MUC5AC_T != "mid") %>%
  survfit(Surv(Observation.Days, Survival) ~ Strat.Score_MUC5AC_T, data = .)

MUC5AC_curve <- ggsurvplot(fit = SurvivalMUC5AC, data = MetaData_Surv, title = "MUC5AC",
           pval = TRUE, 
           conf.int = TRUE, 
           risk.table = TRUE, 
           risk.table.col = "strata",
           surv.median.line = "hv",
           legend.labs = c("low","high"),
           break.x.by= 365,
           ggtheme = theme_classic()
           )
MUC5AC_curve
rm(SurvivalMUC5AC)

SurvivalMUC6 <- MetaData_Surv %>%
  filter(., Strat.Score_MUC6_T != "mid") %>%
  survfit(Surv(Observation.Days, Survival) ~ Strat.Score_MUC6_T, data = .)

MUC6_curve <- ggsurvplot(fit = SurvivalMUC6, data = MetaData_Surv, title = "MUC6",
           pval = TRUE, 
           conf.int = TRUE, 
           risk.table = TRUE, 
           risk.table.col = "strata",
           surv.median.line = "hv",
           legend.labs = c("low","high"),
           break.x.by= 365,
           ggtheme = theme_classic()
           )
MUC6_curve
rm(SurvivalMUC6)

SurvivalMUC2 <- survfit(Surv(Observation.Days, Survival) ~ Strat.Score_MUC2_T, data = MetaData_Surv)

MUC2_curve <- ggsurvplot(fit = SurvivalMUC2, data = MetaData_Surv, title = "MUC2", 
           pval = TRUE, 
           conf.int = TRUE, 
           risk.table = TRUE, 
           risk.table.col = "strata",
           surv.median.line = "hv",
           legend.labs = c("low", "mid","high"),
           break.x.by= 365,
           ggtheme = theme_classic()
           )
MUC2_curve
rm(SurvivalMUC2)

SurvivalMUC4 <- survfit(Surv(Observation.Days, Survival) ~ Strat.Score_MUC4_T, data = MetaData_Surv)

MUC4_curve <- ggsurvplot(fit = SurvivalMUC4, data = MetaData_Surv, title = "MUC4", 
           pval = TRUE, 
           conf.int = TRUE, 
           risk.table = TRUE, 
           risk.table.col = "strata",
           surv.median.line = "hv",
           legend.labs = c("low", "mid","high"),
           break.x.by= 365,
           ggtheme = theme_classic()
           )
MUC4_curve
rm(SurvivalMUC4)

SurvivalMUC13 <- survfit(Surv(Observation.Days, Survival) ~ Strat.Score_MUC13_T, data = MetaData_Surv)

MUC13_curve <- ggsurvplot(fit = SurvivalMUC13, data = MetaData_Surv, title = "MUC13",
           pval = TRUE, 
           conf.int = TRUE, 
           risk.table = TRUE, 
           risk.table.col = "strata",
           surv.median.line = "hv",
           legend.labs = c("low", "mid","high"),
           break.x.by= 365,
           ggtheme = theme_classic()
           )
MUC13_curve
rm(SurvivalMUC13)

surv_gastric <- ggarrange(MUC1_curve$plot, MUC5AC_curve$plot, MUC6_curve$plot, common.legend= TRUE, ncol = 3,legend = "right") 
surv_intestinal <- ggarrange(MUC2_curve$plot, MUC4_curve$plot, MUC13_curve$plot, common.legend = TRUE,ncol = 3, legend = "right" )

Mucin_survivalPlot <- ggarrange(surv_gastric,surv_intestinal, nrow = 2)
Mucin_survivalPlot
rm(MUC1_curve,MUC5AC_curve,MUC6_curve,MUC2_curve,MUC4_curve,MUC13_curve,surv_gastric,surv_intestinal)

##fitting a Cox proportional hazards model
MetaData_Surv$Strat.Score_MUC13_T <- relevel(as.factor(MetaData_Surv$Strat.Score_MUC13_T), ref = "mid")
MetaData_Surv$Strat.Score_MUC1_T <- relevel(as.factor(MetaData_Surv$Strat.Score_MUC1_T), ref = "mid")
MetaData_Surv$Strat.Score_MUC2_T <- relevel(as.factor(MetaData_Surv$Strat.Score_MUC2_T), ref = "mid")
MetaData_Surv$Strat.Score_MUC4_T <- relevel(as.factor(MetaData_Surv$Strat.Score_MUC4_T), ref = "mid")
MetaData_Surv$Strat.Score_MUC5AC_T <- relevel(as.factor(MetaData_Surv$Strat.Score_MUC5AC_T), ref = "mid")
MetaData_Surv$Strat.Score_MUC6_T <- relevel(as.factor(MetaData_Surv$Strat.Score_MUC6_T), ref = "mid")

CoxStrat <- coxph(Surv(Observation.Days, Survival) ~ Age + Gender + Strat.Score_MUC13_T + Strat.Score_MUC1_T + Strat.Score_MUC2_T + Strat.Score_MUC4_T + Strat.Score_MUC5AC_T +  Strat.Score_MUC6_T, data = MetaData_Surv)
summary(CoxStrat)
Forest_coxStrat <- ggforest(CoxStrat)
Forest_coxStrat

#testing cox-proportional hazards model assumptions
test.pheno <- cox.zph(CoxStrat)
test.pheno

ggcoxzph(test.pheno)

ggcoxdiagnostics(CoxStrat,  type = "dfbeta", linear.predictions = FALSE, ggtheme = theme_bw())

ggcoxdiagnostics(CoxStrat,type = "deviance", linear.predictions = FALSE, ggtheme = theme_bw())

rm(CoxStrat,test.pheno)
```

#adding the clinical and expression data to the phyloseq object
```{r}
#reading the clinical data
#manually importing the different metadata 
MetaData_Ill <- MetaData %>%
  drop_na(data = ., "Illumina.ID") %>%
  as.data.frame(.)
rownames(MetaData_Ill) <- MetaData_Ill$Illumina.ID

sample_data(agglom_BA) <- MetaData_Ill
sample_data(agglom_relAb) <- MetaData_Ill
```

