---
title: "3.final_filtering"
author: "Baptiste Oosterlinck"
date: "15-2-2022"
output: html_document
---
Function for plotting the abundances of the samples
```{r}
plot_abundance = function(physeq,title = "", Facet = "Order", Color = "Phylum"){
# Arbitrary subset, based on Phylum, for plotting 
p1f = subset_taxa(physeq, Phylum %in% c("Firmicutes")) 
mphyseq = psmelt(p1f) 
mphyseq <- subset(mphyseq, Abundance > 0) 
ggplot(data = mphyseq, mapping = aes_string(x = "Tissue.Type",y = "Abundance", color = Color, fill = Color)) +
geom_violin(fill = NA) + geom_point(size = 1, alpha = 0.3, position = position_jitter(width = 0.3)) +
facet_wrap(facets = Facet) + scale_y_log10()+ theme(legend.position="none")
}
```

#1. Checking the rarefication curves to assess reading depth
  - all curves should achieve stable plateau
  - samples falling below 3000 final reads are insufficiently sampled

```{r}
track_filter <- as.data.frame(rowSums(otu_table(agglom_BA)))
colnames(track_filter) <- c("prev_filter")
track_filter$ID <- row.names(track_filter)
track_fin$ID <- row.names(track_fin)
track_fin <- full_join(track_fin, track_filter, by = "ID")

rare_samples <- track_fin[track_fin$prev_filter <3000,"ID"]
rareSet <- prune_samples(sample_names(agglom_BA) %in% rare_samples, agglom_BA)
rarecurves <- vegan::rarecurve(otu_table(rareSet),step = 50)
```

##rarefaction filtering and minimum read setting
  Filtering the samples that don't achieve a plateau in the rarefication curves or fall below 3000 final reads
```{r}
#create new phyloseq object without the possitive and negative controls and the samples that don't achieve plateau
agglom_BA2 <- subset_samples(agglom_BA, !sample_names(agglom_relAb) %in% c("pos","neg","G33N","A9N","LiLife00206","G67N","A4N","A25T","G117T"))
agglom_relAb2 <- subset_samples(agglom_relAb, !sample_names(agglom_relAb) %in% c("pos","neg","G33N","A9N","LiLife00206","G67N","A4N","A25T","G117T"))

#create new phyloseq object with only the paired samples and the control samples
nonPaired.PID <- c("171574812","581227 064A40","P00145809","G1988","431014 063A05","P00141446","P00133049","270604 038A63")
agglom_relAb3 <- subset_samples(agglom_relAb2, !Patient.ID %in% nonPaired.PID)
agglom_BA3 <- subset_samples(agglom_BA2, !Patient.ID %in% nonPaired.PID)

#exporting the tracking table of the reads
writexl::write_xlsx(x= track_fin,col_names = TRUE,path = "ReadTracking_IlluminaPool_20220221.xlsx")
```

#Cleaning up the workspace

```{r}
rm(list = setdiff(ls(),c("agglom_BA","agglom_BA2","agglom_BA3","agglom_relAb","agglom_relAb2","agglom_relAb3")))
```
